MCMC Redistricting

### 2020-12-11 paper.Rmd
add paper file

---
title: "MCMC Redistricting"
author: "Dawson Eliasen, Jonathan Olavarria, Henrique Rio"
date: "12/7/2020"
output: beamer_presentation
---

@@

## Redistricting Problem Representation
- Redistricting is represented as a “graph cut” problem
- We can represent a state as a graph $G$
- Each precinct $i$ is a node in $V$
- There is an undirected edge $i \sim j$ connecting nodes $i, j$ if the precincts $i, j$ share a border
- Then, we get voter districts by partitioning the set of nodes $V$
- Since districts are formed by “cutting” edges, we can think of redistricting as a “graph cut” problem

## Monte Carlo Markov Chain (MCMC)
The Metropolis-Hastings algorithm is used to simulate redistricting plans.

@@

## MCMC Example
To simulate redistricting, “perturbation” looks like swapping precincts along existing district borders.

@@

## `redist` Package
- The authors of the paper use a package called `redist` which implements this Metropolis-Hastings algorithm for simulating redistricting
- `redist.mcmc` requires precinct adjacency matrix, population data, and assignments to a valid set of congressional districts
- The authors of the paper shared some code for reproduction, but the code was limited to generating figures and it was not generalizable to different data

## Data
- The paper uses data from the Harvard Election Archive, but those data did not have congressional district assignments for Colorado
- Instead, we used a shapefile assembled by MGGG constructed from
    - Election data from Colorado Secretary of State’s Office
    - 2010 Census data from the U.S. Census Bureau Data API
    - Precinct borders from U.S. Census Bureau TIGER/Line shapefile repository

## Simulation
- Using the `sf` package, we read the shapefiles and calculated a precinct adjacency matrix with a package called `spdep`
- Then, we used `redist.mcmc` to generate 3 MCMC chains of 10,000 samples each - this took a few hours
    - We initially tried 3 MCMC chains of 100,000 samples (this is what the authors of the paper do) - this took 12 hours and crashed my laptop
    - We also tried applying a “compactness” constraint, but it requires providing a distance matrix, which took too long to compute
- Finally, we evaluate bias in the simulations using `redist.segcalc` and a modified version of the code shared by the authors of the paper

@@

```{r, echo=FALSE}
co <- read_sf(dsn='co_precincts', layer='co_precincts')
co <- co %>% arrange(CD116FP)
from <- unique(co$CD116FP)
to <- seq_len(length(unique(co$CD116FP)))
co$CD116FP <- co$CD116FP %>% mapvalues(from, to)

@@

```{r, echo=FALSE}
plot_sim_bias(xax, co_sim_bias)
```

## Minimum Bias Map
```{r, echo=FALSE}
plot_min_bias(co_sim, co_sim_bias, co, xax)
```

>>>>

---
title: "MCMC Redistricting"
author: "Dawson Eliasen, Jonathan Olavarria, Henrique Rio"
date: "12/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

# Motivation

@@

# Methodology

## Redistricting Problem Representation
Redistricting is represented as a “graph cut” problem, where a state is represented as a graph $G$, and each voting precinct $i$ is a node in the set of nodes $V$. There is an undirected edge $i \sim j$ connecting nodes $i, j$ if the precincts $i, j$ are contiguous (share a border). Then, we get voter districts by partitioning the set of nodes $V$. This leaves subgraphs that represent the congressional districts. Since districts are formed by “cutting” edges, we can think of redistricting as a “graph cut” problem. This is how the MCMC methods generate new samples of redistricting plans.

## Metropolis-Hastings
The Metropolis-Hastings algorithm is a Monte Carlo Markov Chain algorithm based on the simple accept-reject algorithm, used to simulate redistricting plans.

@@

## Metropolis-Hastings Iteration Example
To simulate redistricting, “perturbation” looks like swapping precincts along existing district borders. \ref{fig:MCMC example} illustrates the perturbation process for an example sample in a Metropolis-Hastings chain. Perturbation works by identifying nodes that exist on congressional district borders and proposing swaps along the borders. This produces a perturbation of the previous sample which is accepted with acceptance ratio

$$
\alpha(\pi', \mathbf{CP} | \pi) = \min \Big(1,
    \Big(\frac{|B(\mathbf{CP}, \pi)|}{|B(\mathbf{CP}, \pi'|}\Big)^R 
    \frac{F(|B(\mathbf{CP}, \pi)|)(1- q)^{C(\pi', \mathbf{V_{CP}})}}{F(|B(\mathbf{CP}, \pi')|) (1- q)^{C(\pi, \mathbf{V_{CP}})}} \cdot
    \frac{g(\pi')}{g(\pi)} 
\Big)
$$

where $\pi$ is the current sample, $\pi'$ is the perturbed sample, $\mathbf{CP}$ is the set of connected components formed by $\pi$, and $C$ is a graph-cut function.

@@

## `redist` Package
Fifield et. al. use a package called `redist` which implements this Metropolis-Hastings algorithm for simulating redistricting. `redist.mcmc` is a function for generating redistricting samples, and requires precinct adjacency matrix, population data, and assignments to a valid set of congressional districts as function inputs. The authors of the paper shared some code for reproduction, but the code was limited to generating figures and it was not generalizable to different data, so we re-implemented the simulation code to generate redistricting simulations for the state of Colorado.

## Data
Fifield et. al. use Pennsylvania shapefiles and election data from the Harvard Election Archive. However, the Harvard Election Archive Colorado dataset did not include congressional district assignments. Instead, we used a shapefile assembled by MGGG which was constructed from election data from Colorado Secretary of State’s Office, 2010 Census data from the U.S. Census Bureau Data API, and precinct borders from U.S. Census Bureau TIGER/Line shapefile repository.

## Simulation
Using the `sf` package, we read the Colorado shapefiles and calculated a precinct adjacency matrix with a package called `spdep`. Then, we used `redist.mcmc` to generate 3 MCMC chains of 10,000 samples each - this took a few hours. We initially tried 3 MCMC chains of 100,000 samples (this is what the authors of the paper do) - but this took to long and required more computing power than was available. We also tried applying a “compactness” constraint, but it requires providing a distance matrix of the nodes in the graph, which took too long to compute. Finally, we evaluate bias in the simulations using `redist.segcalc` and a modified version of the code shared by the authors of the paper.

# Results

@@

```{r, echo=FALSE, fig.cap="Current congressional districts in Colorado.\\label{fig:map}"}
co <- read_sf(dsn='co_precincts', layer='co_precincts')
co <- co %>% arrange(CD116FP)
from <- unique(co$CD116FP)
to <- seq_len(length(unique(co$CD116FP)))
co$CD116FP <- co$CD116FP %>% mapvalues(from, to)

@@

```{r, echo=FALSE, fig.cap="Redistricting samples, percent of precincts changed vs. bias.\\label{fig:mcmc-summary}"}
plot_sim_bias(xax, co_sim_bias)
```

## Minimum Bias Map
```{r, echo=FALSE, fig.cap="Redistricting sample with minimum bias.\\label{fig:min-bias-map}"}
plot_min_bias(co_sim, co_sim_bias, co, xax)
```
